The persuasion of symbolic elements in government public communications, and their moderation by personal relevance

### 2019-07-16 paper3_relevance and symbols_revision_01.Rmd
Add files via upload

---
title: "The persuasion of symbolic elements in government public communications, and their moderation by personal relevance"
author: '**Saar Alon-Barkat**, PhD candidate; the Hebrew University of Jerusalem;   saar.barkat@mail.huji.ac.il'
date: " "
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
  pdf_document:
    toc: yes
  word_document: default
link-citations: yes
bibliography: phd_paper_2.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
urlcolor: blue
---

@@

To assess the above concern, this study explores the boundaries of governments' persuasion of citizens through symbolic communication. Namely, I seek to examine the conditions under which symbolic elements are likely to be more, or less, effective in shaping (and possibly manipulating) citizens' trust. Building on the social psychology Elaboration Likelihood Model (ELM) [@petty_1986; @petty_2011], I theorize that citizens are less susceptible to symbols, and more inclined to scrutinize the substantive content of communication, the more they perceive the policy issue at hand as having personal relevance for themselves, which enhances their motivation to engage in elaborate processing of information. Employing a survey experiment, I test these hypotheses in the context of the Israeli Environment Protection Ministry, and its policy plan for reducing the air pollution in a specific area (the Haifa-Bay). I operationalize the perceived personal relevance of the policy plan both experimentally, and by comparing between citizens from the polluted Haifa-Bay area and others. The results indicate that even high levels of perceived personal relevance do not attenuate the effect of symbols on citizens' trust, contrary to my expectations. These findings may suggest that the boundaries of the effect of symbolic communication are wider than expected. I discuss the theoretical and normative implications of these findings.

@@

The above-mentioned theoretical expectations are tested via a randomized survey experiment, building on the methodology of @petty_1983. The experiment focuses, empirically, on the case of the Israeli Environmental Protection Ministry (hereafter: EPM), and its policy for reducing the air pollution in the highly polluted Haifa-Bay area. In the following section, I describe the empirical case, data collection, experimental design and operationalization of research variables.

@@

EPM is a national-level bureau responsible for the formulation, coordination, and execution of environmental policy in Israel. Among others, EPM is responsible for monitoring and regulating the air quality across the country. The Haifa-Bay area, located in the north of Israel, has been characterized by relatively high levels of air pollution, mainly due to the Haifa-Bay industrial zone, which is located in proximity of a number of urban residential areas. It is considered the largest industrial zone in the country, with concentration of petrochemical industries, including oil refineries, power plants, and chemical factories. Beginning in Mid 2000s, EPM has been developing and implementing a series of policy programs to monitor and reduce the high levels of pollution specifically in that area [@theministryofenvironmentprotection_2015; @theministryofenvironmentprotection_2015a; @theministryofenvironmentprotection_2019]. While EPM has claimed that its programs have been effective and led to a significant reduction in air-pollution levels, local residents, professionals and other public agencies have remained highly concerned about the issue, and  accordingly it remained high on the public agenda [@statecomptrollerofisrael_2019]. Specifically, the issue received much media and public attention in 2015-2016, following a series of reportings an media publications about the health consequences of the air-pollution in Haifa-Bay. In 2015, the Ministry of Health published statistical data indicating that Haifa-Bay residents experience significantly higher rates of cancer and asthma, compared to other areas of the country. The subsequent media reports specifically highlighted the conclusion that half of the cases of cancer among children in the area can be associated to air-pollution [@elroi_website_2015; @raavad_website_2015]. Thereafter, in 2016, media reporters revealed the interim results of a large epidemiological study, establishing a link between air-pollution in the Haifa-Bay and certain types of cancer, as well as abnormal measurements of newborn babies and growth curves of infants.[^footnote_epidemiological_study] The media reports of that study repeatedly focused on the finding that babies in the Haifa-Bay are born with "smaller heads" [@koriel_website_2016; @rinat_website_2016; @rivlin_website_2016]. Some reports also highlighted that the areas that are geographically closer to the industrial areas are found to be at higher risk for cancer [@bendavid_website_2016]. These publications, and others, caused much panic among the local population, and especially among young parents.

@@

The air-pollution in the Haifa-Bay offers a fortunate case for examining the moderating role of perceived personal relevance. The distinction between those who reside in the polluted area and those who do not entails a significant natural variation in that moderating variable. Importantly, because this policy issue is intangible and scientifically complicated, the variation in personal relevance is not strongly confounded with personal knowledge about the issue, as is normally the case [@petty_1983]. Accordingly, those residing in the polluted Haifa-Bay area are likely to be more motivated to scrutinize a government communication about the air-pollution in their area, yet they are not necessarily more capable of critically evaluating its content. That latter assumption was the main reason for selecting that case.

While comparing those residing in the Haifa-Bay to others enables me to examine the moderating effect of perceived personal relevance, it still has an important limitation - it represents an observational comparison. The differences in personal relevance can be theoretically confounded by unobserved covariates, that cannot be effectively controlled in the analyses. I address this limitation in the research design in three main ways. *First*, in the selection of the research population, I sought to reduce differences in socioeconomic and demographic variables between residents of Haifa-Bay, and their comparison group. Accordingly, I selected the major cities in the Haifa-Bay area (*Haifa*, *Nesher*, *Qiryat-Hayiim*,[^footnote_qiryat-Hayiim] *Qiryat-Bialik*, *Qiryat-Mozkin*, *Qiryat-Ata* and *Qiryat-Yam*), and matched them with citizens from cities with a similar profile and population size at the center of the country (*Natania*, *Petach-Tikva* and *Rishon-Lezion*). The socioeconomic characteristics of these selected cities are summarized in the supplementary appendix (section ##). I also searched relevant EPM documents and news websites for any reporting about high air-pollution levels in the three center cities, to ensure that the two groups would differ in relation to the their perceptions about the relevance air-pollution policies.

*Second*, in addition to the policy regarding the air-pollution in Haifa-Bay, I also included in the experiment a second policy plan that is expected to be perceived by the two groups as equally relevant. For this purpose, I selected a policy plan of EPM that regards *reducing domestic waste and increasing recycling*. The comparison between the two EPM policies is intended to confirm that respondents’ divergent responses to symbolic elements and variation in substantive information in the communication of the first policy, if found, are attributed to differences in perceived personal relevance, and not to other covariates. I henceforth refer to these two policy plans as "air-pollution" and "recycling". *Third*, in addition to the observational comparison between the two areas, I also included an experimental manipulation aimed at enhancing the perceived personal relevance of the air-pollution policy among the Haifa-Bay residents.

@@

To select and recruit residents of these selected cities, I was assisted by an Israeli internet research panel company *iPanel*. The survey experiment was conducted online through Qualtrics Survey Software. The online survey link was sent by iPanel between 28.1-1.2.2018. A total of `r N_0` respondents completed the survey (a response rate of `r round((N_0/5077)*100,1)`%). Thereafter, I filtered out `r N_0-N_1` observations (`r round(((N_0-N_1)/N_0)*100,1)`%) due to multiple entries from the same IP address (n=`r filter_IP`), failure at the instructional manipulation test (n=`r filter_IMC`), submitting the survey in less than 3 or more than 30 minutes (n=`r filter_time`), and respondents under the age of 18 (n=`r filter_age`). These filters are neither associated with the areas nor with the experimental conditions. Following this screening, I ended up with a sizable analytical sample of `r N_1` subjects, of whom `r SVIVA2_01_haifa%>%nrow()` live in the Haifa-bay cities, and `r SVIVA2_01_center%>%nrow()` live in the center cities.[^footnote_unfiltered_sample]

@@

The survey was presented to the participants as an academic study about citizens' attitudes regarding environmental policies in general. After a set of pre-manipulation questions, all subjects, regardless of their area of residence, were randomly assigned to treatment and control conditions of perceived personal relevance (hereafter: relevance manipulation). Subjects in the relevance treatment group were asked three questions about their interest in environmental policy issues, the extent to which they follow these issues in the media, and their preference for more/less government investment in this field. Then, they were asked to describe the most salient environmental issue in their area of residence. These treatment questions were designed to prime those residing in the Haifa-bay area to think about the personal relevance of air-pollution in their area.[^footnote_relevance_manipulation] Subjects in the relevance control group were asked equivalent questions about their field of occupation.

[^footnote_relevance_manipulation]: The first three treatment questions were also designed to prime subjects, more generally, to think about the impact of environmental issues to their own lives. In this paper, I focus on the the moderating effect of the relevance manipulation on Haifa-Bay citizens' responses to the air-pollution policy.

@@

In line with the abovementioned hypotheses, my main focus in this paper is on participants' trust in the air-pollution policy, and the two-way interactions between residence in Haifa-Bay and the symbols and information manipulations of that policy. I also explore these interactions with regard to the recycling policy, and compare them to the air-pollution interactions. Additionally, within the Haifa-Bay sample, I examine the two-way interactions between the relevance manipulation and the symbols and information manipulations. I conducted power estimations for the main effects of the symbols and information manipulations, their conditional effects within each area and the relevant two-way interactions. I conducted these estimations after the data collection and analyses, based on the analytical sample size. The estimated detectable interaction coefficient between areas and the real symbols treatment is 0.27 SDs, and between areas and weak information manipulation is 0.23 SDs. As for the interactions with the relevance manipulation (among Haifa-Bay respondents), I estimate that the detectable coefficients of their interactions with the symbols and information conditions are 0.36 and 0.29 SDs, respectively. The power estimations are reported in details in Section ## of the supplementary appendix.

@@

sd.air <- SVIVA2_01_comb%>%summarise(sd(TRUST_air_INDEX,na.rm=T))%>%max()
sd.air.haifa <- SVIVA2_01_comb %>% filter(AREA_center==0)%>%summarise(sd(TRUST_air_INDEX,na.rm=T))%>%max()
```

@@

The key independent variables are the manipulations of symbols[^footnote_symbols_manipulation_check] and information in the communication, as described above, and their interactions with the perceived personal relevance of the communication. The latter is operationalized observationally (through the distinction between the two areas: Haifa-Bay and center) as well as experimentally (through the abovementioned relevance manipulation).

@@

Living in Haifa-bay area was coded according to respondents’ reporting of residence, work or studying in one of the Haifa-Bay cities. To confirm the empirical assumption that the air-pollution policy (but not the recycling policy) is perceived as more personally relevant by Haifa Bay residents compared with center residents, respondents were asked, after the outcome variable items: "To what extent did you feel that the policy on ["Reducing the air-pollution in the Haifa-Bay" / “reducing waste and increasing recycling”] is personally relevant to you?" (1="very little"; 7="very much"). `r figure.x=figure.x+1`**Figure `r figure.x`** shows the distributions of the latter items across the two areas. As expected, the air-pollution policy is perceived as more personally relevant by Haifa Bay residents compared with center residents  (Cohen's D = `r relevance.areas.d[["estimate"]] %>% round(2)` [`r relevance.areas.d[["conf.int"]][["lower"]] %>% round(2)`,`r relevance.areas.d[["conf.int"]][["upper"]] %>% round(2)`]). By comparison, the recycling policy is perceived by both groups as relatively personally relevant by both groups, with no statistically significant differences between them. Additional support for these differences are provided by respondents' answers to the open question about the most salient environmental issue in their area of residence, which was presented to the relevance treatment group (i.e. before presenting the policy plans). Indeed, `r t1<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow()``r t1` of the `r t2<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1) %>% nrow()``r t2` Haifa-Bay subjects who were asked this question regarded in their responses to the air-pollution generated from the industrial area (`r air.content.percent.haifa <- t1/t2*100``r air.content.percent.haifa %>% round(0)`%).[^footnote_air_relevance_coding] Moreover, many Haifa-Bay respondents explicitly linked the air-pollution to higher morbidity rates in their area. Subjects from center cities regarded to a variety of issues, including recycling, traffic jams and cleanliness, and only `r ((SVIVA2_01_center %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow())/nrow(SVIVA2_01_center)*100) %>% round(0)`% of them mentioned air-pollution as a prominent environmental concern in their area.

@@

[^footnote_air_relevance_coding]: I coded any reference to either air-pollution or the chemical factories in the Haifa-Bay area, as related to air-pollution. To confirm inter-coder reliability, a second coder independently coded 70 observations (Krippendorff’s alpha = `r kripp.alpha(df.air.content.reliability)[["value"]] %>% round(2)`).

@@

As for the relevance manipulation, data suggests that assignment to relevance priming questions did not increase Haifa-Bay respondents' perceptions of personal relevance of the air-pollution policy. In fact, those assigned to the treatment group reported sightly lower level of personal relevance on average, yet these differences are not statistically significant (*t* = `r relevance.exp.t[["statistic"]] %>% round(3)`, *p*=`r round(relevance.exp.t[["p.value"]],3)`). This limitation should be kept in mind when interpreting the results regarding this manipulation. The results of all manipulation and validation checks are available from the supplementary appendix, along with summary statistics and correlation matrix for all variables.

@@

`r table.x=table.x+1`**Tables `r table.x`** compares between the different experimental conditions, within each of the two areas, on demographic variables and pre-manipulation covariates. These comparisons confirm that these random assignments yielded fairly balanced groups.[^footnote_balancing_groups_exceptions] The samples of the two areas are fairly balanced with regard to gender, overall trust in national level government ministries and reported interest in environmental policy issues. Yet, Haifa-Bay respondents are somewhat older, more educated, and reported having more left-wing ideology.

@@

Next, I examine the interactions between the symbols and information manipulation, and the two areas. To reiterate, based on my theoretical hypotheses, I expect that subjects living in the Haifa-Bay, who perceive this issue as personally relevant, would be less affected by the symbolic elements, compared with those living in the Center (H<sub>1</sub>). Likewise, I expect Haifa-bay respondents to be more affected by the differences in the content (strong versus weak information), compared with center respondents (H<sub>2</sub>). `r figure.x=figure.x+1`**Figure `r figure.x`** graphically illustrates the conditional effects of the manipulations, within each of the two areas.[^footnote_raw_means] As for the effect of the symbols manipulation (presented in the left panel), the results show that the symbols had a greater effect on those living in the polluted Haifa-Bay area. Their exposure to the real symbols increased their trust in the policy by `r (cond.symbols.haifa$contrasts %>% tidy() %>% .[1,"estimate"] %>% max(.)/sd.air) %>% abs() %>% round(2)` SDs [`r (t.symbols.no.haifa[["conf.int"]][2]/sd.air) %>% abs() %>% round(2)`-`r (t.symbols.no.haifa[["conf.int"]][1]/sd.air) %>% abs() %>% round(2)`] compared with the no symbols group, and `r (cond.symbols.haifa$contrasts %>% tidy() %>% .[3,"estimate"] %>% max(.)/sd.air) %>% abs() %>% round(2)` SDs [`r (t.symbols.fake.haifa[["conf.int"]][1]/sd.air) %>% abs() %>% round(2)`-`r (t.symbols.fake.haifa[["conf.int"]][2]/sd.air) %>% abs() %>% round(2)`] compared to the fake symbols group. The center participants who saw the real symbols also reported, on average, higher levels of trust in the policy, compared with the two control groups. However, these differences were smaller compared with the effect among the Haifa-Bay subjects, and not sufficiently significant. The effects of the information manipulation on the two groups are presented in the right panel. This comparison shows that the differences between participants' trust in the weak and strong policies were roughly similar in both groups.

@@

These results are at odds with the two theoretical hypotheses. They suggest that that those residing in the polluted area were equally, and perhaps even more, likely to be influenced by symbolic elements in the communication, and also to a lesser extent, by the unfamiliar graphical elements included in the fake symbols condition. Additionally, these analyses do not provide evidence that the more engaged Haifa-Bay were more likely than others to scrutinize the content of the communication, and differentiate between strong and weak arguments.

@@

`r table.x=table.x+1`**Table `r table.x`** replicates the models in **Table `r table.x-1`** for the recycling policy. Overall, these models reveal relatively similar patterns to those found in the air-pollution policy. Notice that Haifa-Bay subjects reported lower trust on that policy as well, which may indicate that they their negative perceptions about the Ministry's actions regarding the air-pollution in their area shapes their general perceptions of the Ministry. The directions of the simple effects and interaction terms remain, while the main effects and conditional effects are smaller and less significant.[^footnote_recycling_model] Consistently with the air-pollution policy, Haifa-Bay subjects were more affected by the symbolic elements, and the respondents from the two areas responded similarly to the differences in the information. The fact that similar patterns are observed across the these policies further weakens the initial assumption regarding the moderating role of personal relevance. It also rules out a possible argument that greater relevance enhances (rather then reduces) the effect of symbols. If this was the case, then there should have been no interaction between symbols and the areas for the recycling policy.

@@

Contrary to my expectations, and those of extant research in social psychology and marketing, I did not find that high levels of perceived personal relevance attenuate the effect of symbols on citizens' trust. Subjects residing in the polluted Haifa-Bay indeed reported that they perceive the policy plan for reducing air-pollution in that area as highly relevant for their lives, compared to those who reside in the center. However, despite these significant differences, Haifa-bay subjects did not pay more attention to the logical quality of the policy plan, and they were equally affected by the symbols in the communication. If anything, the results suggest that the symbols had a greater effect on Haifa-bay subjects. Furthermore, priming subjects to think about the personal relevance of environmental policy, and of the consequences of air-pollution in particular, also failed to moderate the effects of the symbols and information in the communications. These findings are also consistent with the findings of @alonbarkat_2017, where they found that the trust of citizens who recently personally experienced prolonged blackouts (and presumably perceived electricity services as personally relevant) was nonetheless significantly affected by a symbol of the government-owned electricity company.

>>>>

---
title: "The persuasion of symbolic elements in government public communications, and their moderation by personal relevance"
author: '**Saar Alon-Barkat**'
date: " "
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
  pdf_document:
    toc: yes
  word_document: default
link-citations: yes
bibliography: phd_paper_2.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
urlcolor: blue
---

@@

To assess the above concern, this study explores the boundaries of governments' persuasion of citizens through symbolic communication. Namely, I seek to examine the conditions under which symbolic elements are likely to be more, or less, effective in shaping (and possibly manipulating) citizens' trust. Building on the social psychology Elaboration Likelihood Model (ELM) [@petty_1986; @petty_2011], I theorize that citizens are less susceptible to symbols, and more inclined to scrutinize the substantive content of communication, the more they perceive the policy issue at hand as having personal relevance for themselves, which enhances their motivation to engage in elaborate processing of information. Employing a survey experiment, I test these hypotheses in the context of the Israeli Environment Protection Ministry, and its policy plan for reducing the air pollution in a specific area (the Haifa Bay). I operationalize the perceived personal relevance of the policy plan both experimentally, and by comparing between citizens from the polluted Haifa Bay area and others. The results indicate that even high levels of perceived personal relevance do not attenuate the effect of symbols on citizens' trust, contrary to my expectations. These findings may suggest that the boundaries of the effect of symbolic communication are wider than expected. I discuss the theoretical and normative implications of these findings.

@@

The above-mentioned theoretical expectations are tested via a randomized survey experiment, building on the methodology of @petty_1983. The experiment focuses, empirically, on the case of the Israeli Environmental Protection Ministry (hereafter: EPM), and its policy for reducing the air pollution in the highly polluted Haifa Bay area. In the following section, I describe the empirical case, data collection, experimental design and operationalization of research variables.

@@

EPM is a national-level bureau responsible for the formulation, coordination, and execution of environmental policy in Israel. Among others, EPM is responsible for monitoring and regulating the air quality across the country. The Haifa Bay area, located in the north of Israel, has been characterized by relatively high levels of air pollution, mainly due to the Haifa Bay industrial zone, which is located in proximity of a number of urban residential areas. It is considered the largest industrial zone in the country, with concentration of petrochemical industries, including oil refineries, power plants, and chemical factories. Beginning in Mid 2000s, EPM has been developing and implementing a series of policy programs to monitor and reduce the high levels of pollution specifically in that area [@theministryofenvironmentprotection_2015; @theministryofenvironmentprotection_2015a; @theministryofenvironmentprotection_2019]. While EPM has claimed that its programs have been effective and led to a significant reduction in air-pollution levels, local residents, professionals and other public agencies have remained highly concerned about the issue, and  accordingly it remained high on the public agenda [@statecomptrollerofisrael_2019]. Specifically, the issue received much media and public attention in 2015-2016, following a series of reportings an media publications about the health consequences of the air-pollution in Haifa Bay. In 2015, the Ministry of Health published statistical data indicating that Haifa Bay residents experience significantly higher rates of cancer and asthma, compared to other areas of the country. The subsequent media reports specifically highlighted the conclusion that half of the cases of cancer among children in the area can be associated to air-pollution [@elroi_website_2015; @raavad_website_2015]. Thereafter, in 2016, media reporters revealed the interim results of a large epidemiological study, establishing a link between air-pollution in the Haifa Bay and certain types of cancer, as well as abnormal measurements of newborn babies and growth curves of infants.[^footnote_epidemiological_study] The media reports of that study repeatedly focused on the finding that babies in the Haifa Bay are born with "smaller heads" [@koriel_website_2016; @rinat_website_2016; @rivlin_website_2016]. Some reports also highlighted that the areas that are geographically closer to the industrial areas are found to be at higher risk for cancer [@bendavid_website_2016]. These publications, and others, caused much panic among the local population, and especially among young parents.

@@

The air-pollution in the Haifa Bay offers a fortunate case for examining the moderating role of perceived personal relevance. The distinction between those who reside in the polluted area and those who do not entails a significant natural variation in that moderating variable. Importantly, because this policy issue is intangible and scientifically complicated, the variation in personal relevance is not strongly confounded with personal knowledge about the issue, as is normally the case [@petty_1983]. Accordingly, those residing in the polluted Haifa Bay area are likely to be more motivated to scrutinize a government communication about the air-pollution in their area, yet they are not necessarily more capable of critically evaluating its content. That latter assumption was the main reason for selecting that case.

While comparing those residing in the Haifa Bay to others enables me to examine the moderating effect of perceived personal relevance, it still has an important limitation - it represents an observational comparison. The differences in personal relevance can be theoretically confounded by unobserved covariates, that cannot be effectively controlled in the analyses. I address this limitation in the research design in three main ways. *First*, in the selection of the research population, I sought to reduce differences in socioeconomic and demographic variables between residents of Haifa Bay, and their comparison group. Accordingly, I selected the major cities in the Haifa Bay area (*Haifa*, *Nesher*, *Qiryat-Hayiim*,[^footnote_qiryat-Hayiim] *Qiryat-Bialik*, *Qiryat-Mozkin*, *Qiryat-Ata* and *Qiryat-Yam*), and matched them with citizens from cities with a similar profile and population size at the center of the country (*Natania*, *Petach-Tikva* and *Rishon-Lezion*). The socioeconomic characteristics of these selected cities are summarized in the supplementary appendix (section ##). I also searched relevant EPM documents and news websites for any reporting about high air-pollution levels in the three center cities, to ensure that the two groups would differ in relation to the their perceptions about the relevance air-pollution policies.

*Second*, in addition to the policy regarding the air-pollution in Haifa Bay, I also included in the experiment a second policy plan that is expected to be perceived by the two groups as equally relevant. For this purpose, I selected a policy plan of EPM that regards *reducing domestic waste and increasing recycling*. The comparison between the two EPM policies is intended to confirm that respondents’ divergent responses to symbolic elements and variation in substantive information in the communication of the first policy, if found, are attributed to differences in perceived personal relevance, and not to other covariates. I henceforth refer to these two policy plans as "air-pollution" and "recycling". *Third*, in addition to the observational comparison between the two areas, I also included an experimental manipulation aimed at enhancing the perceived personal relevance of the air-pollution policy among the Haifa Bay residents.

@@

To select and recruit residents of these selected cities, I was assisted by an Israeli internet research panel company *iPanel*. The survey experiment was conducted online through Qualtrics Survey Software. The online survey link was sent by iPanel between 28.1-1.2.2018. A total of `r N_0` respondents completed the survey (a response rate of `r round((N_0/5077)*100,1)`%). Thereafter, I filtered out `r N_0-N_1` observations (`r round(((N_0-N_1)/N_0)*100,1)`%) due to multiple entries from the same IP address (n=`r filter_IP`), failure at the instructional manipulation test (n=`r filter_IMC`), submitting the survey in less than 3 or more than 30 minutes (n=`r filter_time`), and respondents under the age of 18 (n=`r filter_age`). These filters are neither associated with the areas nor with the experimental conditions. Following this screening, I ended up with a sizable analytical sample of `r N_1` subjects, of whom `r SVIVA2_01_haifa%>%nrow()` live in the Haifa Bay cities, and `r SVIVA2_01_center%>%nrow()` live in the center cities.[^footnote_unfiltered_sample]

@@

The survey was presented to the participants as an academic study about citizens' attitudes regarding environmental policies in general. After a set of pre-manipulation questions, all subjects, regardless of their area of residence, were randomly assigned to treatment and control conditions of perceived personal relevance (hereafter: relevance manipulation). Subjects in the relevance treatment group were asked three questions about their interest in environmental policy issues, the extent to which they follow these issues in the media, and their preference for more/less government investment in this field. Then, they were asked to describe the most salient environmental issue in their area of residence. These treatment questions were designed to prime those residing in the Haifa Bay area to think about the personal relevance of air-pollution in their area.[^footnote_relevance_manipulation] Subjects in the relevance control group were asked equivalent questions about their field of occupation.

[^footnote_relevance_manipulation]: The first three treatment questions were also designed to prime subjects, more generally, to think about the impact of environmental issues to their own lives. In this paper, I focus on the the moderating effect of the relevance manipulation on Haifa Bay citizens' responses to the air-pollution policy.

@@

In line with the abovementioned hypotheses, my main focus in this paper is on participants' trust in the air-pollution policy, and the two-way interactions between residence in Haifa Bay and the symbols and information manipulations of that policy. I also explore these interactions with regard to the recycling policy, and compare them to the air-pollution interactions. Additionally, within the Haifa Bay sample, I examine the two-way interactions between the relevance manipulation and the symbols and information manipulations. I conducted power estimations for the main effects of the symbols and information manipulations, their conditional effects within each area and the relevant two-way interactions. I conducted these estimations after the data collection and analyses, based on the analytical sample size. The estimated detectable interaction coefficient between areas and the real symbols treatment is 0.27 SDs, and between areas and weak information manipulation is 0.23 SDs. As for the interactions with the relevance manipulation (among Haifa Bay respondents), I estimate that the detectable coefficients of their interactions with the symbols and information conditions are 0.36 and 0.29 SDs, respectively. The power estimations are reported in details in Section ## of the supplementary appendix.

@@

sd.air <- SVIVA2_01_comb%>%summarise(sd(TRUST_air_INDEX,na.rm=T))%>%max()
sd.air.haifa <- SVIVA2_01_comb %>% filter(AREA_center==0)%>%summarise(sd(TRUST_air_INDEX,na.rm=T))%>%max()

sd.waste <- SVIVA2_01_comb%>%summarise(sd(TRUST_waste_INDEX,na.rm=T))%>%max()

@@

The key independent variables are the manipulations of symbols[^footnote_symbols_manipulation_check] and information in the communication, as described above, and their interactions with the perceived personal relevance of the communication. The latter is operationalized observationally (through the distinction between the two areas: Haifa Bay and center) as well as experimentally (through the abovementioned relevance manipulation).

@@

Living in Haifa Bay area was coded according to respondents’ reporting of residence, work or studying in one of the Haifa Bay cities. To confirm the empirical assumption that the air-pollution policy (but not the recycling policy) is perceived as more personally relevant by Haifa Bay residents compared with center residents, respondents were asked, after the outcome variable items: "To what extent did you feel that the policy on ["Reducing the air-pollution in the Haifa Bay" / “reducing waste and increasing recycling”] is personally relevant to you?" (1="very little"; 7="very much"). `r figure.x=figure.x+1`**Figure `r figure.x`** shows the distributions of the latter items across the two areas. As expected, the air-pollution policy is perceived as more personally relevant by Haifa Bay residents compared with center residents  (Cohen's D = `r relevance.areas.d[["estimate"]] %>% round(2)` [`r relevance.areas.d[["conf.int"]][["lower"]] %>% round(2)`,`r relevance.areas.d[["conf.int"]][["upper"]] %>% round(2)`]). By comparison, the recycling policy is perceived by both groups as relatively personally relevant by both groups, with no statistically significant differences between them. Additional support for these differences are provided by respondents' answers to the open question about the most salient environmental issue in their area of residence, which was presented to the relevance treatment group (i.e. before presenting the policy plans). Indeed, `r t1<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow()``r t1` of the `r t2<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1) %>% nrow()``r t2` Haifa Bay subjects who were asked this question regarded in their responses to the air-pollution generated from the industrial area (`r air.content.percent.haifa <- t1/t2*100``r air.content.percent.haifa %>% round(0)`%).[^footnote_air_relevance_coding] Moreover, many Haifa Bay respondents explicitly linked the air-pollution to higher morbidity rates in their area. Subjects from center cities regarded to a variety of issues, including recycling, traffic jams and cleanliness, and only `r ((SVIVA2_01_center %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow())/nrow(SVIVA2_01_center)*100) %>% round(0)`% of them mentioned air-pollution as a prominent environmental concern in their area.

@@

[^footnote_air_relevance_coding]: I coded any reference to either air-pollution or the chemical factories in the Haifa Bay area, as related to air-pollution. To confirm inter-coder reliability, a second coder independently coded 70 observations (Krippendorff’s alpha = `r kripp.alpha(df.air.content.reliability)[["value"]] %>% round(2)`).

@@

As for the relevance manipulation, data suggests that assignment to relevance priming questions did not increase Haifa Bay respondents' perceptions of personal relevance of the air-pollution policy. In fact, those assigned to the treatment group reported sightly lower level of personal relevance on average, yet these differences are not statistically significant (*t* = `r relevance.exp.t[["statistic"]] %>% round(3)`, *p*=`r round(relevance.exp.t[["p.value"]],3)`). This limitation should be kept in mind when interpreting the results regarding this manipulation. The results of all manipulation and validation checks are available from the supplementary appendix, along with summary statistics and correlation matrix for all variables.

@@

`r table.x=table.x+1`**Tables `r table.x`** compares between the different experimental conditions, within each of the two areas, on demographic variables and pre-manipulation covariates. These comparisons confirm that these random assignments yielded fairly balanced groups.[^footnote_balancing_groups_exceptions] The samples of the two areas are fairly balanced with regard to gender, overall trust in national level government ministries and reported interest in environmental policy issues. Yet, Haifa Bay respondents are somewhat older, more educated, and reported having more left-wing ideology.

@@

Next, I examine the interactions between the symbols and information manipulation, and the two areas. To reiterate, based on my theoretical hypotheses, I expect that subjects living in the Haifa Bay, who perceive this issue as personally relevant, would be less affected by the symbolic elements, compared with those living in the Center (H<sub>1</sub>). Likewise, I expect Haifa Bay respondents to be more affected by the differences in the content (strong versus weak information), compared with center respondents (H<sub>2</sub>). `r figure.x=figure.x+1`**Figure `r figure.x`** graphically illustrates the conditional effects of the manipulations, within each of the two areas.[^footnote_raw_means] As for the effect of the symbols manipulation (presented in the left panel), the results show that the symbols had a greater effect on those living in the polluted Haifa Bay area. Their exposure to the real symbols increased their trust in the policy by `r (cond.symbols.haifa$contrasts %>% tidy() %>% .[1,"estimate"] %>% max(.)/sd.air) %>% abs() %>% round(2)` SDs [`r (t.symbols.no.haifa[["conf.int"]][2]/sd.air) %>% abs() %>% round(2)`-`r (t.symbols.no.haifa[["conf.int"]][1]/sd.air) %>% abs() %>% round(2)`] compared with the no symbols group, and `r (cond.symbols.haifa$contrasts %>% tidy() %>% .[3,"estimate"] %>% max(.)/sd.air) %>% abs() %>% round(2)` SDs [`r (t.symbols.fake.haifa[["conf.int"]][1]/sd.air) %>% abs() %>% round(2)`-`r (t.symbols.fake.haifa[["conf.int"]][2]/sd.air) %>% abs() %>% round(2)`] compared to the fake symbols group. The center participants who saw the real symbols also reported, on average, higher levels of trust in the policy, compared with the two control groups. However, these differences were smaller compared with the effect among the Haifa Bay subjects, and not sufficiently significant. The effects of the information manipulation on the two groups are presented in the right panel. This comparison shows that the differences between participants' trust in the weak and strong policies were roughly similar in both groups.

@@

These results are at odds with the two theoretical hypotheses. They suggest that that those residing in the polluted area were equally, and perhaps even more, likely to be influenced by symbolic elements in the communication, and also to a lesser extent, by the unfamiliar graphical elements included in the fake symbols condition. Additionally, these analyses do not provide evidence that the more engaged Haifa Bay were more likely than others to scrutinize the content of the communication, and differentiate between strong and weak arguments.

@@

`r table.x=table.x+1`**Table `r table.x`** replicates the models in **Table `r table.x-1`** for the recycling policy (*Mean* = `r SVIVA2_01_comb%>%summarise(mean(TRUST_waste_INDEX,na.rm=T))%>%max() %>% round(2)`, *SD* = `r SVIVA2_01_comb%>%summarise(sd(TRUST_waste_INDEX,na.rm=T))%>%max() %>% round(2)`, *Median* = `r SVIVA2_01_comb%>%summarise(median(TRUST_waste_INDEX,na.rm=T))%>%max() %>% round(2)`). Overall, these models reveal relatively similar patterns to those found in the air-pollution policy. Notice that Haifa Bay subjects reported lower trust on that policy as well, which may indicate that they their negative perceptions about the Ministry's actions regarding the air-pollution in their area shapes their general perceptions of the Ministry. The directions of the simple effects and interaction terms remain, while the main effects and conditional effects are smaller and less significant.[^footnote_recycling_model] Consistently with the air-pollution policy, Haifa Bay subjects were more affected by the symbolic elements, and the respondents from the two areas responded similarly to the differences in the information. The fact that similar patterns are observed across the these policies further weakens the initial assumption regarding the moderating role of personal relevance. It also rules out a possible argument that greater relevance enhances (rather then reduces) the effect of symbols. If this was the case, then there should have been no interaction between symbols and the areas for the recycling policy.

@@

Contrary to my expectations, and those of extant research in social psychology and marketing, I did not find that high levels of perceived personal relevance attenuate the effect of symbols on citizens' trust. Subjects residing in the polluted Haifa Bay indeed reported that they perceive the policy plan for reducing air-pollution in that area as highly relevant for their lives, compared to those who reside in the center. However, despite these significant differences, Haifa Bay subjects did not pay more attention to the logical quality of the policy plan, and they were equally affected by the symbols in the communication. If anything, the results suggest that the symbols had a greater effect on Haifa Bay subjects. Furthermore, priming subjects to think about the personal relevance of environmental policy, and of the consequences of air-pollution in particular, also failed to moderate the effects of the symbols and information in the communications. These findings are also consistent with the findings of @alonbarkat_2017, where they found that the trust of citizens who recently personally experienced prolonged blackouts (and presumably perceived electricity services as personally relevant) was nonetheless significantly affected by a symbol of the government-owned electricity company.
